Data Preparation, Integration and Cleaning

### 2018-05-10 scripts/3_data_prep.Rmd
Complete prep re-write.  Start viz script

---
title: "Data Preparation, Integration and Cleaning"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

@@

### Introduction

@@

### Preliminary Commands

@@

@@

  source(file.path(code_path, 'functions', 'abb_Functions.R'))
  source(file.path(code_path, 'functions', 'geom_sf.R'))

@@

### Constants and Parameters

@@

 # Set all colors
  col_df <- data.frame(names=c('rausch', 'hackberry', 'kazan', 'babu', 'lima', 
                               'beach', 'ebisu', 'tirol', 'foggy', 'hof'),
                       rgb = c(rgb(250,88,99, maxColorValue=255), 
                               rgb(139,14,82, maxColorValue=255),
                               rgb(0,117,140, maxColorValue=255), 
                               rgb(4,211,191, maxColorValue=255),
                               rgb(77,226,110, maxColorValue=255), 
                               rgb(252,179,14, maxColorValue=255),
                               rgb(249,177,139, maxColorValue=255), 
                               rgb(191,166,100, maxColorValue=255),
                               rgb(156,161,155, maxColorValue=255), 
                               rgb(86,94,97, maxColorValue=255)))

 # Set specific colors
  str_col <- col_df$rgb[1]
  ltr_col <- col_df$rgb[3]
  sm_col <- col_df$rgb[c(2,4,5,8,6)]

@@

  sub_sf <- sf::st_read(file.path(data_path, 'geographic', 'melbSuburbs.shp')) %>%
    st_transform(crs=4283)
  sa1_sf <- sf::st_read(file.path(data_path, 'geographic', 'sa1s.shp')) %>%
    st_transform(crs=4283)
  beach_sf <- sf::st_read(file.path(data_path, 'geographic','portPhillipBeach.shp')) %>%
    st_transform(crs=4283)

@@

We then clip by time.  For the short-term data we will consider any Airbnb property that had at least 6 months worth of listing time on the portal during the 12-month study period

@@

  # Short Term data

  str_tdf <- str_tdf %>%
    dplyr::mutate(type = 'Other')
  str_tdf$type[str_tdf$property_type %in% 
               c('Apartment', 'Condominium', 'Loft')] <- 'Apartment'
  str_tdf$type[str_tdf$property_type %in% c('House', 'Townhouse', 'Villa')] <- 'House'

  # Long Term data

  ltr_tdf <- ltr_tdf %>%
    dplyr::mutate(type = 'Other')
  ltr_tdf$type[ltr_tdf$property_type %in% c('Duplex', 'House',
                                             'Terrace', 'Townhouse', 'Villa')] <- 'House'
  ltr_tdf$type[ltr_tdf$property_type %in% c('Studio', 'Unit')] <- 'Apartment'

@@

  # Create long term rental plot

@@

   ggplot(str_tdf, aes(x=listing_type)) + 
      geom_bar(color=str_col, fill=str_col) +
      xlab('\nListing Type\n') +
      ggtitle('\nShort Term Rentals by Listing Type\n')  +
      ylab('# of Properties\n') +
     theme(plot.title = element_text(hjust = 0.5))

@@

We start by looking at bedrooms. We begin by removing all properties that are missing bedrooms information or have a likely data error in this fields (more than 14 bedrooms). We then plot the distribution of diffeltr bedrooms counts.  From this point onwards we bifurcate our analysis between houses and apartments as the markets (prices, rates, and properties) differ between the two property types.

@@

  # Remove missing and data errors

@@

### Location

@@

  # Set colors
  tm_shape(sub_sf) +
    tm_polygons('suburbDesi',
                title='Suburb Designation',
                border.col='white',
                palette=c('gray10', 'gray40', 'gray70'))

@@

  # Add suburb designation and remove those that are outside

   str_sf <- st_join(str_sf, 
                     sub_sf %>% 
                      dplyr::select(suburb = NAME_2006,
                                    sub_des = suburbDesi,
                                    geometry)) %>%
   dplyr::filter(!is.na(sub_des))

   ltr_sf <- st_join(ltr_sf, 
                     sub_sf %>% 
                       dplyr::select(suburb = NAME_2006,
                                     sub_des = suburbDesi,
                                     geometry)) %>%
   dplyr::filter(!is.na(sub_des))

@@

    sub_str <- str_sf %>% 
      dplyr::select(suburb, -geometry) %>% 
      as.tibble() %>%
      dplyr::group_by(suburb) %>% 
      dplyr::summarize(count=n()) %>%
      dplyr::filter(count >= 3 & 
                      suburb != 'Maddingley')

@@

    sub_all <- sub_ltr %>%
      dplyr::inner_join(sub_str, by='suburb')

    str_sf <- str_sf %>%
      dplyr::filter(suburb %in% sub_all$suburb)

@@

  all_sf <- rbind(str_sf %>% 
                    rename(uniq_id = property_id),
                  ltr_sf %>%
                    rename(uniq_id = addr_id))
  ssub_sf <- sub_sf %>% filter(NAME_2006 %in% sub_all$suburb)

  all_ch <- st_buffer(st_convex_hull(st_union(ssub_sf)), .001)
  within_ch <- st_contains(all_ch, sub_sf$geometry)
  subs_within <- sub_sf[unlist(within_ch), ]
  study_area <- st_union(subs_within)
  plot(study_area)

@@

  ltr_tdf %>% 
    dplyr::mutate(geometry = NULL) %>%
    as.tibble()

  str_tdf %>% 
    dplyr::mutate(geometry = NULL) %>%
    as.tibble()

@@

Finally we add the spatial data back onto the tabular data (filtering out those outside of the study area using an inner join).

  str_tdf <- str_tdf %>%
    dplyr::inner_join(str_sf, 
                      by='property_id')

  ltr_tdf <- ltr_tdf %>%
    dplyr::inner_join(ltr_sf,
                      by='addr_id')

@@

  ltr_price <- ggplot(ltr_tdf %>% tidyr::unnest(), aes(x=price)) +
    geom_density(fill=ltr_col, alpha=.8) +
    ylab('') +
    xlab('Weekly Rental') +
    ggtitle('Long Term Rentals') +
    scale_x_continuous(limits=c(0, 1500)) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    theme(plot.title = element_text(hjust = 0.5))

  # Nightly Rate plot
  str_price <- ggplot(str_tdf, aes(x=med_rate)) +
    geom_density(fill=str_col, alpha=.8) +
    ylab('') +
    xlab('Nightly Rate') +
    ggtitle('Short Term Rentals') +
    scale_x_continuous(limits=c(0, 600)) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    theme(plot.title = element_text(hjust = 0.5))

@@

```{r ltprice_fltr2s, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}
# 
#   png(file=paste0(fig.path, 'dataclean_rates.png'), width = 480, height = 280, 
#       bg='transparent')
#     
#     ggMultiPlots(ltr.plot, rate.plot, cols=2)
#   
#   dev.off()
#   
#   save(ltr.plot, file=paste0(fig.path, 'rentplot.RData'))
#   save(rate.plot, file=paste0(fig.path, 'rateplot.RData'))
# 
#   
```

@@

   # Add together for plotting
   rev_df <- bind_rows(str_tdf %>% 
                         dplyr::select(type, gross_revenue) %>%
                         dplyr::mutate(tenure = 'str'),
                       ltr_tdf %>%
                         dplyr::select(type, gross_revenue) %>%
                         dplyr::mutate(tenure = 'ltr'))

   revdens <- ggplot(rev_df, 
                     aes(x=gross_revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=c(ltr_col, str_col)) +
    scale_color_manual(values=c(ltr_col, str_col)) +
    ggtitle('Annual Revenues by Tenure Type') +
    xlab('Annual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20,000', '$40,000', '$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000))

   # Add together for plotting potential revenues
   prev_df <- bind_rows(str_tdf %>% 
                         dplyr::select(type, gross_revenue=pot_revenue) %>%
                         dplyr::mutate(tenure = 'str'),
                       ltr_tdf %>%
                         tidyr::unnest() %>%
                         dplyr::select(type, gross_revenue) %>%
                         dplyr::mutate(tenure = 'ltr'))

   prevdens <- ggplot(prev_df, 
                     aes(x=gross_revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=c(ltr_col, str_col)) +
    scale_color_manual(values=c(ltr_col, str_col)) +
    ggtitle('Annual Revenues by Tenure Type') +
    xlab('Annual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20,000', '$40,000', '$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000))

@@

The location of the short and long term rentals, colored by submarket, are shown below

```{r locationmaps.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

 # Set submarket colors
   # Make map
  tm_shape(sub_sf) +
    tm_polygons() + 
  tm_shape(study_area) + 
    tm_polygons(col='dodgerblue3') +
  tm_shape(ltr_sf) +
    tm_symbols(size=.05, col='black', border.alpha=0, alpha=.02, shape=15)

  tm_shape(sub_sf) +
    tm_polygons() + 
  tm_shape(study_area) + 
    tm_polygons(col='dodgerblue3') +
  tm_shape(str_sf) +
    tm_symbols(size=.05, col='black', border.alpha=0, alpha=.2, shape=15)

  str_map <- ggplot() +
     geom_polygon(data=ssub_sf, aes(x=long, y=lat, group=group), fill='white', color='grey70') +
#     geom_point(data=str_df, aes(x=longitude, y=latitude, color=geo.mrkt), size=.2,
#                alpha=.2) +
#     scale_color_manual(values=sm_col) +
#     scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
#     scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
#     xlab('') + ylab('')+
#     theme(legend.position = 'bottom',
#           legend.title = element_blank(),
#           plot.title = element_text(hjust = 0.5),
#           panel.grid.major = element_blank(),
#           panel.grid.minor = element_blank(),
#           panel.background = element_rect(fill = "white"),
#           axis.text.x = element_blank(),
#           axis.text.y = element_blank(),
#           axis.ticks = element_blank(),
#           axis.title.x=element_blank(),
#           axis.title.y=element_blank(),
#           panel.border = element_rect(colour = "black", fill=NA, size=1)) +
#     guides(colour = guide_legend(override.aes = list(size=3,
#                                                      alpha=1))) + 
#     ggtitle('Short-Term Rental Locations')+
#     theme(plot.title = element_text(hjust = 0.5))
#   
# ```
# 
# ```{r ltr.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}
# 
#   # Make map
# 
#   ltr.map <- ggplot() +
#     geom_polygon(data=subs, aes(x=long, y=lat, group=group), fill='white', color='grey70') +
#     geom_point(data=ltr_df, aes(x=longitude, y=latitude, color=geo.mrkt), size=.1,
#                alpha=.1) +
#     scale_color_manual(values=sm_col) +
#     scale_x_continuous(limits=c(min(subs$long), max(subs$long))) +
#     scale_y_continuous(limits=c(min(subs$lat), max(subs$lat))) +
#     xlab('') + ylab('')+
#     theme(legend.position = 'bottom',
#           legend.title = element_blank(),
#           plot.title = element_text(hjust = 0.5),
#           panel.grid.major = element_blank(),
#           panel.grid.minor = element_blank(),
#           panel.background = element_rect(fill = "white"),
#           axis.text.x = element_blank(),
#           axis.text.y = element_blank(),
#           axis.ticks = element_blank(),
#           axis.title.x=element_blank(),
#           axis.title.y=element_blank(),
#           panel.border = element_rect(colour = "black", fill=NA, size=1)) +
#     guides(colour = guide_legend(override.aes = list(size=3,
#                                                      alpha=1))) + 
#     ggtitle('Long-Term Rental Locations')+
#     theme(plot.title = element_text(hjust = 0.5))
#   
# ```
# 
# ```{r loc.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.width=9}
#   
#   png(file=paste0(fig.path, 'str_locmap.png'), width = 680, height = 590, 
#       bg='transparent')
#      str.map
#   dev.off()
#   
#   png(file=paste0(fig.path, 'ltr_locmap.png'), width = 680, height = 590, 
#       bg='transparent')
#      ltr.map
#   dev.off()
# 
#   ggMultiPlots(str.map, ltr.map, cols=2)
#   
#   save(str.map, file=paste0(fig.path, 'strMap.RData'))
#   save(ltr.map, file=paste0(fig.path, 'ltrMap.RData'))
# 
# 
# ```
# 
# The chart below break down the relative frequency of short and long term rental properties across the five submarkets. 
# 
```{r smtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.width=9}

   # Make Airbnb submarket map

   strtype.map <- ggplot() +
     geom_bar(data=str_tdf, aes(x=geo_mrkt, fill=geo_mrkt)) +
     scale_fill_manual(values=abb_col[c(1,3,5,7,9)]) +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'bottom',
           legend.title = element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank()) +
     #ggtitle('Short Term Property Type\nBy Submarket')+
     theme(plot.title = element_text(hjust = 0.5))

   # Make Long term rental map

   ltrtype.map <- ggplot() +
     geom_bar(data=ltr_tdf, aes(x=geo_mrkt, fill=geo_mrkt)) +
     scale_fill_manual(values=abb_col[c(1,3,5,7,9)]) +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'bottom',
           legend.title = element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank()) +
     #ggtitle('Long Term Property Type\nBy Submarket')+
     theme(plot.title = element_text(hjust = 0.5))

```
# 
# ```{r loctype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.width=9}
# 
#   png(file=paste0(fig.path, 'submplot.png'), width = 680, height = 590, 
#       bg='transparent')
#     ggMultiPlots(strtype.map, ltrtype.map, cols=2)
#   dev.off()
#     
#   ggMultiPlots(strtype.map, ltrtype.map, cols=2)
#   
#   save(strtype.map, file=paste0(fig.path, 'strsubplot.RData'))
#   save(ltrtype.map, file=paste0(fig.path, 'ltrsubplot.RData'))
# 
#

@@

   save(str_df, ltr_df, daily_df, listing.data, subs, abb_col, 
#        beach.shp, core.suburbs, exch.rate, ltr_col, str_col, sa1s.shp, 
#        suburbs.shp, study.suburbs, clean.count, daily.run.total, list.run.total,
#        ltr.run.total, str.run.total,
#        file=paste0(data_path, 'prepared/prep_workspace_', data_name, '.RData'))
# 
# ```N

>>>>

---
title: "Data Preparation, Integration and Cleaning"
output:
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

@@

## Introduction

@@

## Preliminary Commands

@@

Then we set the paths to the data

@@

  source(file.path(code_path, 'functions', 'abb_Functions.R'))

@@

## Constants and Parameters

@@

 # Set all colors
  col_df <- data.frame(names=c('rausch', 'hackberry', 'kazan', 'babu', 'lima', 
                               'beach', 'ebisu', 'tirol', 'foggy', 'hof'),
                       rgb = c(rgb(250, 88, 99, maxColorValue=255), 
                               rgb(139, 14, 82, maxColorValue=255),
                               rgb(0, 117, 140, maxColorValue=255), 
                               rgb(4, 211, 191, maxColorValue=255),
                               rgb(77, 226, 110, maxColorValue=255), 
                               rgb(252, 179, 14, maxColorValue=255),
                               rgb(249, 177, 139, maxColorValue=255), 
                               rgb(191, 166, 100, maxColorValue=255),
                               rgb(156, 161, 155, maxColorValue=255), 
                               rgb(86, 94, 97, maxColorValue=255)))

 # Set specific colors
  str_col <- col_df$rgb[1]
  ltr_col <- col_df$rgb[3]
  sm_col <- col_df$rgb[c(2, 4, 5, 8, 6)]

@@

  suburb_sf <- sf::st_read(file.path(data_path, 'geographic', 'melbSuburbs.shp')) %>%
    st_transform(crs=4283)
  sa1_sf <- sf::st_read(file.path(data_path, 'geographic', 'sa1s.shp')) %>%
    st_transform(crs=4283)
  beach_sf <- sf::st_read(file.path(data_path, 'geographic','portPhillipBeach.shp')) %>%
    st_transform(crs=4283)

@@

We then clip by time.  For the short-term data we will consider any Airbnb property that had at least 6 months worth of listing time on the portal during the 12-month study period.  We also filter the long term data to include only observations with one year prior to the decision date.

@@

  # Short Term data  
  str_tdf <- str_tdf %>%
    dplyr::mutate(type = 'Other')
  str_tdf$type[str_tdf$property_type %in% 
               c('Apartment', 'Condominium', 'Loft')] <- 'Apartment'
  str_tdf$type[str_tdf$property_type %in% c('House', 'Townhouse', 'Villa')] <- 'House'

  # Long Term data  
  ltr_tdf <- ltr_tdf %>%
    dplyr::mutate(type = 'Other')
  ltr_tdf$type[ltr_tdf$property_type %in% c('Duplex', 'House',
                                             'Terrace', 'Townhouse', 'Villa')] <- 'House'
  ltr_tdf$type[ltr_tdf$property_type %in% c('Studio', 'Unit')] <- 'Apartment'

@@

@@

   # Plot by listing type
   ggplot(str_tdf, aes(x=listing_type)) + 
      geom_bar(color=str_col, fill=str_col) +
      xlab('\nListing Type\n') +
      ggtitle('\nShort Term Rentals by Listing Type\n')  +
      ylab('# of Properties\n') +
     theme(plot.title = element_text(hjust = 0.5))

@@

We start by looking at bedrooms. We begin by removing all properties that are missing bedrooms information or have a likely data error in this fields (more than 10 bedrooms). We then plot the distribution of different bedrooms counts.  From this point onwards we bifurcate our analysis between houses and apartments as the markets (prices, rates, and properties) differ between the two property types.

@@

@@

@@

  # Set colors
  tm_shape(suburb_sf) +
    tm_polygons('suburbDesi',
                title='Suburb Designation',
                border.col='white',
                palette=c('gray10', 'gray40', 'gray70'))

@@

  # Add suburb designation and remove those that are outside   
   str_sf <- st_join(str_sf, 
                     suburb_sf %>% 
                      dplyr::select(suburb = NAME_2006,
                                    sub_des = suburbDesi,
                                    geometry)) %>%
   dplyr::filter(!is.na(sub_des))

   ltr_sf <- st_join(ltr_sf, 
                     suburb_sf %>% 
                       dplyr::select(suburb = NAME_2006,
                                     sub_des = suburbDesi,
                                     geometry)) %>%
   dplyr::filter(!is.na(sub_des))

@@

    # Count those with at least 3 transactions
    sub_str <- str_sf %>% 
      dplyr::select(suburb, -geometry) %>% 
      as.tibble() %>%
      dplyr::group_by(suburb) %>% 
      dplyr::summarize(count=n()) %>%
      dplyr::filter(count >= 3 & 
                      suburb != 'Maddingley')

@@

    # Join to create unified list
    sub_all <- sub_ltr %>%
      dplyr::inner_join(sub_str, by='suburb')

    # Remove those with insufficient observations
    str_sf <- str_sf %>%
      dplyr::filter(suburb %in% sub_all$suburb)

@@

  # Get ids of study suburbs
  sa_ids <- suburb_sf %>%
    dplyr::filter(NAME_2006 %in% sub_all$suburb) %>%
    sf::st_union() %>%
    sf::st_convex_hull() %>%
    sf::st_buffer(., .001) %>%
    st_contains(., suburb_sf$geometry) %>%
    unlist()

  # Study Area
  studyarea_sf <- suburb_sf %>% 
    slice(sa_ids) %>%
    st_union()
  plot(studyarea_sf)

  # Create Indicators in main sf fle
  suburb_sf <- suburb_sf %>%
    dplyr::select(name = NAME_2006, class = suburbDesi, geometry) %>%
    dplyr::mutate(in_study = if_else(name %in% sub_all$suburb, 1, 0),
                  in_studyarea = 0)
  suburb_sf$in_studyarea[sa_ids] <- 1

@@

Finally we add the spatial data back onto the tabular data and filtering out those outside of the study area using an inner join. We then have to remove the geometry field as it causes issues with future analyses (for now)

  str_tdf <- str_tdf %>%
    dplyr::inner_join(str_sf, 
                      by='property_id') %>% 
     dplyr::mutate(geometry = NULL) %>%
     as.tibble()

  ltr_tdf <- ltr_tdf %>%
    dplyr::inner_join(ltr_sf,
                      by='addr_id') %>% 
     dplyr::mutate(geometry = NULL) %>%
     as.tibble()

@@

@@

  ggplot(ltr_tdf %>% tidyr::unnest(), aes(x=price)) +
    geom_density(fill=ltr_col, alpha=.8) +
    ylab('') +
    xlab('Weekly Rental') +
    ggtitle('Long Term Rentals') +
    scale_x_continuous(limits=c(0, 1500)) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    theme(plot.title = element_text(hjust = 0.5))

  # Nightly Rate plot
  ggplot(str_tdf, aes(x=med_rate)) +
    geom_density(fill=str_col, alpha=.8) +
    ylab('') +
    xlab('Nightly Rate') +
    ggtitle('Short Term Rentals') +
    scale_x_continuous(limits=c(0, 600)) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    theme(plot.title = element_text(hjust = 0.5))

@@

@@

@@

@@

   save(str_tdf, ltr_tdf, str_sf, ltr_sf, studyarea_sf, suburb_sf, beach_sf, sa1_sf,
        file=paste0(data_path, '/prepared/prepared_data.RData'))
